---
title: "masscor package"
author: "Cristhian Paredes"
date: "`r Sys.Date()`"
output: 
  knitr:::html_vignette:
    toc: true
    fig_caption: yes
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{masscor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib  
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# Render html vignetes by using devtools::document(roclets = "vignette")
# Render also pdf vignetes by using rmarkdown::render("vignettes/labsimplex.Rmd", "all")
```

The R package `masscor` provides functions, classes and methods to support mass measurements using non automatic balances as described in @euramet. The new classes are objects that can store the calibration information for balances and mass standards. Those objects can be used to convert balance readings to both conventional mass and mass, and to perform routine balance verification (by using the normalized error function). Air buoyancy correction factors are calculated using local air density that can be calculated using measured environmental conditions and applying one of several models available in the package. The uncertainty of (corrected) mass measurements can also be evaluated allowing to further asses the suitability of a given mass measurement.

This package contains example data sets with real and hypothetical calibration data for some balances, a single mass standard and a mass standards kit. This vignette is intended as a user manual for the `masscor` package.

**Disclaimer:** The Instituto Nacional de Metrologia de Colombia (INM) has published this package to ease some calculations involved in mass metrology applications. We endeavor to update it on a regular basis, but cannot guarantee the accuracy of all implemented methods. Any reference to commercial measurement instruments does not imply any approval, endorsement or recommendation by the INM.

## Installation and use

The last released version of the package `masscor` can be installed from CRAN by running:

```{r,  eval = FALSE}
install.packages("masscor")
```

The development version of masscor can be installed from [GitHub](https://github.com/Crparedes/masscor "masscor GitHub Repository") using `install_github()` function from `devtools` package [@devtools]:

```{r,  eval = FALSE}
devtools::install_github("Crparedes/masscor", build_vignettes = TRUE)
```

To use the functions in `masscor`, the package must be loaded into R environment after the installation by running:

```{r setup}
library(masscor)
```

## Balance calibration certificates using the `masscor` package

The calibration of a balance involve tests to determine the repeatability and the error of indications in several points spread along the measurement interval of the balance, and the effect of applying the load in eccentric positions of the balance pan [@euramet].

The repeatability test consist in taking several balance readings for the same object under similar conditions to asses short time variability. This process can be repeated with other objects of different mass to evaluate the repeatability at different balance loads. The errors of the indications are obtained by comparing balance readings to the masses of mass standards that are traceable to internationally recognized references (traceable to the International System of Units). The eccentricity effect is determined by the variation in balance readings for the same object when intentionally placed at different positions of the weighing pan.

The calibration certificate of a balance will typically include information about the balance division scale which is usually also indicated somewhere in the body of the balance (readability), a table of repeated indications for one or more objects (repeatability), a table with indication errors and associated uncertainties for several mass readings in the measurement interval of the balance, and a table with the differences in the readings for the same object located at several positions in the balance pan (eccentricity). 

The indication errors can later be used to correct a given balance reading to a conventional mass value with metrological traceability (see `convMass()`). The uncertainties of the indication errors provide information to estimate the uncertainty in the conventional mass due to this correction.

The package `masscor` uses objects of class `calibCert` to store the information of balance calibration certificates. This class of objects are generated by the function `calibCert()`. Mandatory arguments for this function are the balance division scale, the results of the indication error test, the results of repeatability test, and the results of the eccentricity test:

* Balance division scale is provided at the parameter`d`. The corresponding units can be indicated at the parameter `d.units` (default value is `'mg'`).
* The results of the indication error test are provided as a data frame at the parameter `indError`. The data frame must have three columns containing nominal masses, indication errors and associated uncertainties for at least two mass standards. The corresponding units can be indicated at the parameter `indError.units` as a character vector of length three (one element for nominal masses, indication errors and associated uncertainties, respectively. Default values are `c('g', 'mg', 'mg')`). The uncertainties can correspond to expanded uncertainties in wich case parameter `expanded` is set to `TRUE` (the default) and a coverage factor can be indicated at the parameter `k` (default `k = 2`).
* The repeatability test results are provided at the parameter `rep` and can either be a vector of length two with the nominal mass (mean) and standard deviation of the balance readings under repeatability conditions for the same object, or in case the repeatability test is performed at more than one point, a data frame with the nominal masses (in the first column) and standard deviations (in the second column) of the balance readings under repeatability conditions for each object. The corresponding units can be indicated at the parameter `rep.units` as a vector of length two (one element for the nominal mass and for the standard deviation, respectively. Default values are `c('g', 'mg')`).  
* The eccentricity test results is provided at the parameter `eccen` as a single value indicating the maximal difference in the balance readings for the same object placed at different positions. The corresponding units can be indicated as a character at the parameter `eccen.units` (default value is `'mg'`).

The default values for the parameters are those typical for an analytical balance calibration certificate. Significant miscalculations can occur down the pipe if the units in calibration certificate differ and are not indicated when creating the object of class `calibCert`.

Suppose the simplest case in which an analytic balance (scale division of 0.1\ mg) is calibrated using only two (traceable) mass standards of nominal masses 5 and 20\ g. The indication errors are -0.2 $\pm$ 0.200 and -0.1 $\pm$ 0.205\ mg, respectively. The standard deviation of six readings under repeatability conditions for the mass standard of 20\ g is 0.02\ mg, and the maximal difference in the eccentricity test was 0.1\ mg. This information is enough to produce an object of class `calibCert` that can later be used to correct a balance reading in the interval between 5 and 20\ g and to later estimate mass measurement uncertainties. A balance identification string can be provided indicating, for example, the brand, type and location of the balance.

```{r calibCert1}
Balance.1_Lab.317 <- calibCert(balanceID = '(Brand) Analytic balance lab 317', 
                               d = 0.1, d.units = 'mg',
                               indError = data.frame(c(5, 20), c(-0.2, -0.1), c(0.200, 0.201)) # miligrams
                               indError.units = c('g', 'mg', 'mg'),
                               rep = c(20, 0.02), rep.units = c('g', 'mg'),
                               eccen = 0.1, eccen.units = 'mg')
```

The package provides methods to `print` and `plot` the information contained in objects of class `calibCert`:
```{r calibCert2}
print(Balance.1_Lab.317)
plot(Balance.1_Lab.317)
```

The objects of class `calibCert` must be created when the balance undergoes its first calibration and must be updated each time the balance undergoes recalibration. The best way to share the `calibCert` objects among the people that uses the balance is by saving the `calibCert` object as a `.RData` file using the R function `save()`. When required, the `calibCert` object can be loaded into the R Environment by using the function `load()`. 

```{r calibCert3}
save(Balance.1_Lab.317, file = 'certificateBalance.1_Lab.317.RData')
load(file = 'certificateBalance.1_Lab.317.RData')
```
The directory to/from which the `.RData` file is going to be saved/loaded must be kept in mind. See `help(getwd)` for further details.

Additional information of the calibration certificate can be set in the optional parameters of the function `calibCert()` (run `help(calibCert)` for details). Some balance calibration certificates provided as example data sets in the package illustrate the purpose of those optional parameters.
```{r calibCert4}
data(MT.XPE.204)
print(MT.XPE.204)
plot(MT.XPE.204)
```

## Conventional mass corrections

We live and realize mass measurements in air. The air causes the objects to float and this buoyancy affects the measurement of their masses. The real mass of an object (or simply, its mass) would be that determined in complete vacuum or that corrected by the air buoyancy effect. Mass air buoyancy corrections are easy to apply if both the densities of the object whose mass is being measured and that of the air causing it to float are known or properly estimated (see Section Mass Air Buoyancy Correction). As long as this correction is usually very small and the complete information to perform it is not always known, a common practice is to report the mass of objects under arbitrarily agreed conditions, leading to the conventional mass of the object.

The conventional mass of an object is defined as the mass $m_c$ of a mass standard that balances this body under conventionally chosen conditions: at a temperature $t_{ref} = 20^o\text{C}$, with mass standards of density $\rho_c=8000~\text{kg m}^{-3}$, in normal air of density $\rho_0=1.2~\text{kg m}^{-3}$ [@OIML_D28_2004].

The balance calibration information is used to convert a balance reading to conventional mass. The indication error reported for each mass standard in the balance calibration certificate is the opposite of the correction needed for the given balance at that load. The mass correction that must be applied to a given balance reading is obtained by interpolating the value into the straight line defined by the two closest calibration points. This technique known as bracketing corrects the balance reading and provides an uncertainty estimate for the conventional mass obtained due to the correction.

The `masscor` package uses the calibration information contained in an object of class `calibCert` to obtain a conventional mass correction by using the function `convMass()`.

Suppose we use the balance whose calibration certificate is stored in the example data set `MT.XPE.204` provided with the package. For a balance reading of 211.1342\ g, the (corrected) conventional mass is obtained by running:

```{r convMass1}
data(MT.XPE.204)
convMass(reading = 211.1342, units = 'g', calibCert = MT.XPE.204)
```

### Conventional mass uncertainty
The uncertainty in the conventional mass of an object includes both the uncertainty from the indication error correction and the uncertainty in the balance reading which in turn considers the uncertainty due to balance readability, related to the balance division scale and the uncertainty arising from the lack of repeatability.

Uncertainty from the indication error correction by bracketing corresponds to the combination of the uncertainties in the two indication errors used in the interpolation:$$u_{error}=\sqrt{u_{low}^2 + u_{high}^2}$$%
where $u_{low}$ is the uncertainty in the indication error correction for the lower point and $u_{high}$ is the uncertainty in the indication error correction for the higher point used in the interpolation.

Uncertainty in the balance reading combines the lack of repeatability obtained as the standard deviation from measures of the same object's mass, and the effect of the readability obtained from the balance division scale by assuming a triangular probability density function: $$u_{read}=\sqrt{sd^2+\frac{d^2}{6}}$$%
where $sd$ is the measurements standard deviation and $d$ is the balance division scale.

The balance reading and the indication error uncertainties are combined as usual:
$$u_{mass}=\sqrt{u_{error}^2 + u_{read}^2}$$

**preguntar si esto que sigue está bien** 
Mass measurements usually involve taring the balance to perform a mass measurement by substracting the zero. In these terms the uncertainty estimate should include these rest would be $$u_{mass}=\sqrt{u_{error}^2 + 2\cdot u_{read}^2}$$
**Confirmar esto anterior y modificar las funciones segun corresponda...**

The main function to asses the uncertainty in a conventional mass value using `masscor` package is `uncertMassConv()`. This function internally calls the functions `uncertErrorCorr()` and `uncertReading()` that calculate the uncertainties of the indication error correction and balance reading, respectively.

Following with the previous example, assuming the repeated measurements standard deviation is 0.03\ mg, the uncertainty in the conventional mass is:

```{r convMass2}
uncertConvMass(reading = 211.1342, units = 'g', repValues = 0.03e-3, 
               calibCert = MT.XPE.204, tare = TRUE)
``` 
In some cases the experimenter is willing to give up some readability in order to improve the time a mass measurement can take. In those events balances are used with a division scale bigger from the division scale stated in the calibration certificate. The uncertainty in the balance reading must consider the change in the balance division scale so the used division scale can be provided to the functions `uncertMassConv()` and `uncertReading()` as an optional parameter. 

If the measurement from the previous example is performed with a division scale of 1\ mg instead of the default of 0.1\ mg, the conventional mass uncertainty is:

```{r convMass3}
uncertConvMass(reading = 211.134, units = 'g', d = 1e-3, repValues = 0.03e-3, 
               calibCert = MT.XPE.204, tare = TRUE)
```
## Air buoyancy correction
As mentioned in previous section, when measuring the mass of an object in air the object experiences buoyancy and its mass is affected. This effect is usually very small and can be safely ignored most of the times. When the correction is needed, the mass conventional mass of the object being measured can be corrected by multiplying it by its mass air buoyancy correction (MABC) factor defined as $$MABC=\frac{1-\frac{\rho_{air}}{\rho_{weights}}}{1-\frac{\rho_{air}}{\rho_{sample}}}$$%
where $\rho_{air}$ is the local air density (around 1.2\ $kg~m^{-3}$), $\rho_{weights}$ is the density of the mass standards used during calibration (most of the times the mass standards are made of stainless steel with a density of 8000\ $kg~m^{-3}$), and $\rho_{sample}$ is the density of the object.

The function `MABC()` calculates mass air buoyancy correction factors using the above formula. The densities must be provided all in same units.
```{r MABC1}
MABC(rho_s = 2.115, rho_w = 8, rho_air = 0.001199314) # g/mL
MABC(rho_s = 2115, rho_w = 8000, rho_air = 1.199314) # g/mL
``` 
The mass of the object whose conventional mass is 

### Air density models and associated uncertainty

### Uncertainty of air bouyancy correction

## Additional functions in the package

### Data for mass standards and mass standards kits

### Routine balance verification: normalized error
