---
title: "masscor package"
author: "Cristhian Paredes"
date: "`r Sys.Date()`"
output: 
  knitr:::html_vignette:
    toc: true
    fig_caption: yes
  pdf_document:
    highlight: null
    number_sections: yes
vignette: >
  %\VignetteIndexEntry{masscor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: bibliography.bib  
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The R package `masscor` provides functions, classes and methods to support mass measurements using digital balances. 
The classes contain mainly 

## Introduction

Mass measurements involve...

## Installation

The last released version of the package `masscor` can be installed from CRAN by running:

```{r,  eval = FALSE}
install.packages("masscor")
```

The development version of masscor can be installed from [GitHub](https://github.com/Crparedes/masscor "masscor GitHub Repository") using `install_github()` function from `devtools` package [@devtools]:

```{r,  eval = FALSE}
devtools::install_github("Crparedes/masscor", build_vignettes = TRUE)
```

To load `masscor` package into R environment run

```{r setup}
library(masscor)
```

## `masscor` functions

### Creating, exporting and importing the data of a balance calibration certificate

The calibration certificate of a balance typically will provide a table with the indication error 
and associated uncertainty

```{r calibCert1}
massSTD  <- c(0.01, 0.5, 1, 10, 20, 50, 100, 120, 150, 200, 220)  ## [g]
indError <- c(0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, -0.2, -0.2) ## [mg]
uncert   <- c(0.1, 0.1, 0.1, 0.1, 0.1, 0.2, 0.2, 0.3, 0.3, 0.4, 0.5) ## [mg]
d <- 0.1 ## [mg]
traceability <- 'Set of weights class E2. Certificate number 1473 D-K 17296, 2019-05-10.'


MT.XPE.204 <- calibCert(balanceID = 'MT XPE 204', serial = 'B403223982',
                        massSTD = massSTD,
                        indError = indError, uncert = uncert, d = d,
                        units = c('g', 'mg', 'mg', 'mg'),
                        classSTD = 'E2', traceability = traceability,
                        Temp = c(17.4, 17.9), ## [deg.C]
                        p = c(750.4, 751.0), ## [hPa]
                        h = c(70.5, 71.4), ## [%]
                        unitsENV = c('deg.C', 'hPa', '%'),
                        institution = 'Instituto Nacional de Metrologia de Colombia',
                        date = '2021-03-18')
```

The information of the ...

```{r calibCert2}
print(MT.XPE.205)
plot(MT.XPE.205)
```

The object of class `calibCert` created corresponds to the calibration certificate provided in the package...

### Conventional mass correction

The conventional mass of an object is the

#### Conventional mass uncertainty

### Air buoyancy correction

#### Air density models and associated uncertainty

#### Uncertainty of air bouyancy correction

## Additional functions in the package

### Data for mass standards and mass standards kits

### Routine balance verification: normalized error
